---
title: "<center> Non-linear Modeling </center>"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
---
<center> <h3> Sunah Park </center>

This markdown file is created by Sunah Park for extended lab exercises in the book _An Introduction to Statistical Learning with Applications in R_ [(ISLR)](https://www-bcf.usc.edu/~gareth/ISL/ISLR%20First%20Printing.pdf). 


* * *

### Setup for code chunks
```{r mychunksetup, echo=TRUE, include=TRUE} 
rm(list=ls())
# default r markdown global options in document
knitr::opts_chunk$set(
   ########## Text results ##########
    echo=TRUE, 
    warning=FALSE, # to preserve warnings in the output 
    error=FALSE, # to preserve errors in the output
    message=FALSE, # to preserve messages
    strip.white=TRUE, # to remove the white lines in the beginning or end of a source chunk in the output 

    ########## Cache ##########
    cache=TRUE,
   
    ########## Plots ##########
    fig.path="", # prefix to be used for figure filenames
    fig.width=8,
    fig.height=6,
    dpi=200
)
```



* * * 
```{r lib, echo=TRUE, include=TRUE} 
library(ISLR)
library(splines) # bs() function
library(ggplot2)
library(gam) # gam() function
library(akima) # two-dimensional surface
```


```{r df, echo=TRUE, include=TRUE} 
head(Wage,3)
Wage<-na.omit(Wage)
```

* * *
Linear models are relatively simple to describe and implement, and have advantages over other approaches in terms of interpretation and inference. However, standard linear regression can have significant limitations in terms of predictive power. Some simple extensions of linear models are examined here below. 

## __1. Polynomial Regression__
Polynomial regression extends the linear model by adding extra predictors, obtained by raising each of the original predictors to a power. This approach provides a simple way to provide a non-linear fit to a data.
```{r poly, echo=TRUE, include=TRUE} 
fit<-lm(wage~poly(age,4), data=Wage)
coef(summary(fit))

fit2<-lm(wage~poly(age,4, raw=T), data=Wage)
coef(summary(fit2))

fit3<-lm(wage~age+I(age^2)+I(age^3)+I(age^4), data=Wage)
coef(summary(fit3))

fit4<-lm(wage~cbind(age, age^2, age^3, age^4), data=Wage)
coef(summary(fit4))

agelims<-range(Wage$age)
age.grid<-seq(from=agelims[1], to=agelims[2])
preds<-predict(fit, newdata=list(age=age.grid), se=TRUE)
se.bands<-cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)

preds2<-predict(fit2, newdata=list(age=age.grid), se=TRUE)
max(abs(preds$fit-preds2$fit))
```


```{r polyplot, echo=TRUE, include=TRUE} 
plot(Wage$age, Wage$wage, xlim=agelims,col="grey",pch=19)
lines(age.grid, preds$fit, col="red")
matlines(age.grid, se.bands, col="blue", lty=2)
```

### Decision on the degree of polynomial
```{r anova, echo=TRUE, include=TRUE} 
fit.1<-lm(wage~age, data=Wage)
fit.2<-lm(wage~poly(age,2), data=Wage)
fit.3<-lm(wage~poly(age,3), data=Wage)
fit.4<-lm(wage~poly(age,4), data=Wage)
fit.5<-lm(wage~poly(age,5), data=Wage)

anova(fit.1, fit.2, fit.3, fit.4, fit.5)
```

We perform the ANOVA(Analysis of Variance) function in order to test the null hypothesis that a model M1 is sufficient to explain the data against the alternative hypotehsis that a more complex model M2 is required. In order to use the anova() function, M1 and M2 must be __nested models__: the predictors in M1 must be a subset of the predictors in M2. 

The p-value comparing the linear Model 1 to the quadratic Model 2 is essentially zero, indicating that a linear fit is not sufficient. Similarly the p-value comparing the quadratic Model 2 to the cubic Model 3 is very low (0.0017), so the quadratic fit is also insufficient. The p-value comparing the cubic and degree-4 polynomials, Model 3 and Model 4, is approximately 5% while the degree-5 polynomial Model 5 seems unnecessary because its p-value is 0.37. Hence, either a cubic or a quartic polynomial appear to provide a reasonable fit to the data, but lower- or higher-order models are not justified.

In this case, instead of using the anova() function, we could have obtained these p-values more succinctly by exploiting the fact that poly() creates orthogonal polynomials. Note that the p-values are the same, and in fact the square of the t-statistics are squal to the F-statistics from the anova() function.
```{r polyn, echo=TRUE, include=TRUE} 
coef(summary(fit.5))
```

However, the ANOVA method works whether or not we used orthogonal polynomials; it also works when we have other terms in the model as well. 
```{r anova2, echo=TRUE, include=TRUE} 
fit.1<-lm(wage~education+age, data=Wage)
fit.2<-lm(wage~education+poly(age,2), data=Wage)
fit.3<-lm(wage~education+poly(age,3), data=Wage)
anova(fit.1, fit.2, fit.3)
```

As an alternative to using hypothesis tests and ANOVA, we could choose the polynomial degree using __cross-validation__(Chapter 5). 


Next we consider the task of predicting whether an individual earns more than $250,000 per year. I() was used to create the binary response variable. 
```{r glm, echo=TRUE, include=TRUE} 
fit<-glm(I(wage>250)~poly(age,4),data=Wage, family=binomial) # polynomial logistic regression
preds<-predict(fit, newdata=list(age=age.grid), se=TRUE)

pfit<-exp(preds$fit)/(1+exp(preds$fit))
se.bands.logit<-cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
se.bands<-exp(se.bands.logit)/(1+exp(se.bands.logit))
preds<-predict(fit, newdata=list(age=age.grid), type="response", se=T)

plot(Wage$age, I(Wage$wage>250), xlim=agelims, type="n", ylim=c(0,0.2))
points(Wage$age, I((Wage$wage>250)/5), cex=.5, col="grey",pch=19)
lines(age.grid, pfit, col="red")
matlines(age.grid, se.bands, col="blue",lty=2)
```

We have drawn the age values corresponding to the observations with wage values above 250 as gray marks on the top of the plot, and those with wage values below 250 are shown as gray marks on the bottom of the plot. 


* * *
## __2. Step function__
Step functions cut the range of a variable into K distinct regions in order to produce a qualitative variable. This has the effect of fitting a piecewise constant function.
```{r step, echo=TRUE, include=TRUE} 
table(cut(Wage$age,5)) # cut() for step function
fit<-lm(wage~cut(age,5), data=Wage)
preds<-predict(fit, newdata=list(age=age.grid), se=TRUE)
se.bands<-cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)

coef(summary(fit))

plot(Wage$age, Wage$wage, xlim=agelims,pch=19, col="grey")
lines(age.grid, preds$fit, col="red")
matlines(age.grid, se.bands, col="blue",lty=2)
```
cut() function autimatically picks the cutpoints at 30.4, 42.8,55.2 and 67.6. The function cut() returns an ordered categorical variable; the lm() function then creates a set of dummy variables for use in the regression. 


* * *
## __3. Regression Splines__
Regression splines are more flexible than polynomials and step functions, and in fact are an extension of the two. They involve dividing the range of X into K distinct regions. Within each region, a polynomial function is fit to the data. However, these polynomials are constrained so that they join smoothly at the region boundaries, or knots. Provided that the interval is divided into enough regions, this can produce an extremely flexible fit. 
```{r reg-splines, echo=TRUE, include=TRUE} 
fit<-lm(wage~bs(age, knots=c(25,40,60)), data=Wage) # bs: Generate the B-spline basis matrix for a polynomial spline
summary(fit)
pred<-predict(fit, newdata=list(age=age.grid), se=TRUE)
se.bands<-cbind(pred$fit+2*pred$se.fit, pred$fit-2*pred$se.fit)
plot(Wage$age, Wage$wage, col="grey",pch=19)
lines(age.grid, pred$fit, col="red")
matlines(age.grid, se.bands, col="blue", lty=2)
```

Here we have prespecified knots at ages 25, 40 and 60. This produces a spline with six basis functions (A cubic spline with three knots has seven degrees of freedom; these degress of freedom are used up by an intercept, plus six basis functions). We could also use the df option to produce a spline with knots at uniform quantiles of the data.
```{r reg-splines2, echo=TRUE, include=TRUE} 
dim(bs(Wage$age, knots=c(25,40,60))) # bs(): generate the B-spline basis matrix for a polynomial spline
dim(bs(Wage$age, df=6))
attr(bs(Wage$age,df=6), "knots")
```

At df=6, R chooses knots at ages 33.8, 42.0 and 51.0, which correspond to the 25th, 50th and 75th percentiles of age. The function bs() also has a degree argument, so we can fit splines of any degree, rather than the default degree of 3. 

In order to instead fit a natural spline, we use the ns() function. 
```{r ns, echo=TRUE, include=TRUE} 
fit2<-lm(wage~ns(age,df=4), data=Wage)
pred2<-predict(fit2, newdata=list(age=age.grid), se=T)

se.bands<-cbind(pred2$fit+2*pred2$se.fit, pred2$fit-2*pred2$se.fit)
plot(Wage$age, Wage$wage, col="grey",pch=19)
lines(age.grid, pred2$fit, col="red")
matlines(age.grid, se.bands, col="blue", lty=2)
```

As with the bs() function, we could instead specify the knots direclty using the knots option.

* * *
## __4. Smoothing splines__
Smoothing splines are similar to regression splines, but arise in a slightly different situation. Smoothing splines result from minimizing a residual sum of squares criterion subject to a smoothness penalty. 

In order to fit a smoothing spline, we use the smooth.spline() function. 

```{r smooth, echo=TRUE, include=TRUE} 
fit<-smooth.spline(Wage$age, Wage$wage, df=16)
fit2<-smooth.spline(Wage$age, Wage$wage, cv=TRUE)
fit2$df


plot(Wage$age, Wage$wage, xlim=agelims, col="grey", pch=19)
lines(fit, col="red")
lines(fit2, col="blue")
legend("topright", legend=c("16 DF", "6.8 DF"), col=c("red","blue"), lty=1)
```

In the first call of smooth.spline() we specified df=16. The function then determines which value of lambda leads to 16 df. In the second call to smooth.spline() we select the smoothness level by cross-validation; this results in a value of lambda that yields 6.8 df. 

* * *
## __5. Local regression__
Local regression is similar to splines, but differs in an important way. The regions are allowed to overlap, and indeed they do so in a very smooth way.
In order to perform local regression, we use the loess() function.
```{r loess, echo=TRUE, include=TRUE} 
fit<-loess(wage~age, span=.2, data=Wage)
fit2<-loess(wage~age, span=.5, data=Wage)

plot(Wage$age, Wage$wage, xlim=agelims, col="grey",pch=19)
lines(age.grid, predict(fit, data.frame(age=age.grid)), col="red")
lines(age.grid, predict(fit2, data.frame(age=age.grid)),col="blue")
legend("topright", legend=c("span=0.2", "span=0.5"), col=c("red","blue"), lty=1)
```

We performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20% or 50% of the observations. The larger the span, the smoother the fit. 

* * *
## __6. Generalized additive models (GAM)__
GAM allow us to extend the methods above to deal with multiple predictors. 
We fit a GAM to predict wage using natural spline functions of year and age, treating education as a qualitative predictor using gam() function from gam library. 
```{r gam, echo=TRUE, include=TRUE} 
gam1<-lm(wage~ns(year,4)+ns(age,5)+education, data=Wage)
gam.m3<-gam(wage~s(year,df=4)+s(age,df=5)+education, data=Wage)
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE)

par(mfrow=c(1,3))
plot.Gam(gam1, se=TRUE)
```
The s() function is used to indicate that we would like to use a smoothing spline. We specify that the function of year should have 4 degrees of freedom, and that the function of age will have 5 degress of freedom. Since education is qualitative, we leave it as is, and it is converted into four dummy variables. 

Even though gam1 is not of class gam but rather of class lm, we can still use plot.gam() on it. We have to use plot.gam() rather than the generic plot() function.
 
The function of year looks rather linear. We can perform a series of ANOVA tests in order to determine which of these three models is best: a GAM that excludes year(m1), a GAM that uses a linear function of year(m2), or a GAM that uses a spline function of year(m3).
```{r gam2, echo=TRUE, include=TRUE} 
gam.m1<-gam(wage~s(age,5)+education, data=Wage)
gam.m2<-gam(wage~year+s(age,5)+education, data=Wage)
gam.m3<-gam(wage~s(year,4)+s(age,5)+education, data=Wage)
anova(gam.m1, gam.m2, gam.m3, test="F")
```

We find that there is compelling evidence that a GAM with a linear function of year(m2) is beter than a GAM that does not include year at all(m1) (p-value of 0.00014). However, there is no evidence that a non-linear function of year is needed (p-value=0.349). In other words, based on the results of this ANOVA, M2 is preferred.

```{r gam-p, echo=TRUE, include=TRUE} 
summary(gam.m3)
```
The p-values for year and age (from Anova for Nonparametric Effects) correspond to a null hypothesis of a linear relationship versus the alternative of a non-linear relationship. The large p-value for year reinforces our conclusion from the ANOVA test that a linear function is adequate for this term. However, there is very clear evidence that a non-linear term is required for age. 

```{r gam-pred, echo=TRUE, include=TRUE} 
pred<-predict(gam.m2, newdata=Wage)
# We can also use local regression fits as building blocks in a GAM, using the lo() function
gam.lo<-gam(wage~s(year, df=4)+lo(age, span=0.7)+education, data=Wage)
par(mfrow=c(1,3))
plot.Gam(gam.lo, se=TRUE)


## Local regression using lo()
gam.lo.i<-gam(wage~lo(year, age, span=0.5)+education, data=Wage) # fits a two-term model, in which the first term is an interaction between year and age, fit by a local regression surface
par(mfrow=c(1,2))
plot(gam.lo.i)


## Logistic Regression GAM using I()
gam.lr<-gam(I(wage>250)~year+s(age, df=5)+education, family=binomial, data=Wage) 
par(mfrow=c(1,3))
plot(gam.lr, se=T)

table(Wage$education, I(Wage$wage>250)) # We see that there are no high earners in the < HS category
gam.lr.s<-gam(I(wage>250)~year+s(age,df=5)+education, family=binomial, data=Wage, 
              subset=(education!="1. < HS Grad")) # Fit a logistic regression GAM using all but education 1. < HS Grad -> This provides more sensible results
par(mfrow=c(1,3))
plot(gam.lr.s, se=T) 

```


